Modelling the emergence of whisker barrels

### 2019-12-17 paper/briefpaper/brief.tex
Change sigma to eta, because we have dsigma in Eq 4

\dropcap{S}patial patterns in neural connectivity provide clues about the constraints under which brains evolve and develop \citep{purves_iterated_1992}. Perhaps the most distinctive pattern can be found in the rodent barrel cortex \cite{woolsey_structural_1970}. The barrels are large cortical columns, identifiable from birth in layer 4 of primary somatosensory cortex as dense clusters of thalamocortical axons, which are enclosed by borders a few neurons thick from postnatal day 3 \citep{erzurumlu_development_2012}. In the plane tangential to the cortical surface the barrels constitute a somatotopic map of the whiskers, with cells of adjacent barrels responding most strongly and quickly to deflection of adjacent whiskers \citep{armstrong-james_flow_1992}. Barrel patterning reflects subcortical whisker maps comprising cell aggregates called barrelettes in the brainstem and barreloids in the thalamus \cite{van_der_loos_barreloids_1976}.

Barrel formation requires afferent input from whisker stimulation and thalamic calcium waves \citep{anton-bolanos_prenatal_2019}, and depends on a complex network of axon guidance molecules such as ephrin-A5 and A7 and adhesion molecules such as cadherin-6 and 8 \citep{miller_epha7-ephrin-a5_2006}.
This network is orchestrated by interactions between morphogens Fgf8 and Fgf17 and transcription factors Emx2, Pax6, Sp8, and Coup-tf1 \citep{shimogori_fibroblast_2005,bishop_regulation_2000}, which are expressed in gradients that mark orthogonal axes and can be manipulated to stretch, shrink, shift, and even duplicate barrels \cite{assimacopoulos_fibroblast_2012}.

@@

\caption{\textbf{A.} Cortical whisker barrels from \cite{zheng_signal_2001}
  annotated with lines joining interstices into a putative Voronoi diagram
  ($\delta=0.022$). \textbf{B.}  Top: Stochastic initial condition for $a_i$
  with $i=1$ (barrel `$\alpha$'). The roll-off towards the boundary (thin line)
  ensures the boundary conditions are satisfied. Bottom: Guidance molecules
  $\rho_1(\mathbf{x})$ and $\rho_2(\mathbf{x})$ in the simulated
  cortex. $\rho_1(\mathbf{x})$ varies along the anterior-posterior axis;
  $\rho_2(\mathbf{x})$ is expressed along the lateral/medial axis, analogous
  to ephrin A5 \citep{vanderhaeghen_mapping_2000}. \textbf{C.}  Barreloid
  centres for mystacial whiskers taken from \cite{haidarliu_size_2001} are
  shown as black circles. The dashed grey line shows the boundary of the
  barreloid field. It is assumed that there are two linearly increasing,
  orthogonal signalling molecules in the thalamus, such that the location of
  each barreloid within the graded thalamic signal field can differentiate the
  efferent axons, with the result that axons from each barreloid interact to
  varying degrees with the cortical signalling gradients $\rho_1$ and
  $\rho_2$. Blue (green) arrows show, for each barreloid, the magnitude (and
  sign) of the interaction that its thalamocortical projection will have with
  $\rho_1$ ($\rho_2$). The mappings from barreloid location to interaction
  strength are linear. \textbf{D.}  Barrel cortex simulation with parameters
  $\alpha=3$, $\beta=20$, $k=3$, $D=0.2$, $\gamma\in\pm 2$ and
  $\delta{t}=0.0001$. Colours indicate $\arg\max_i(c_i)$ for 41 TC
  projections.  \textbf{E.}  The Honda Voronoi measure, $\delta$, (red) and a
  pattern fitting metric based on the distance, $d_i$, between simulated and
  experimental barrel centroids and the absolute difference in their areas,
  $\Delta A_i$ given by $\sigma = \sum_i^N d_i^2 \times \sum_i^N \Delta A_i$
  (black) plotted as a function of time for the simulations shown in
  D. \textbf{F.}  Guidance molecules in the simulated Fgf8 misexpression
  experiment. $\rho_1$ is mirrored over the anterior-posterior axis. The
  simulation indicates that the mirroring of the barrel field can occur as a
  result of the mirroring of guidance molecule 1.}
\label{fig:main}
\end{figure*}

>>>>

\dropcap{S}patial patterns in neural connectivity provide clues about the constraints under which brains evolve and develop \citep{purves_iterated_1992}. Perhaps the most distinctive pattern can be found in the rodent barrel cortex \cite{woolsey_structural_1970}. The barrels are large cortical columns, identifiable from birth in layer 4 of primary somatosensory cortex as dense clusters of thalamocortical axons, which are enclosed by borders a few neurons thick from postnatal day 3 \citep{erzurumlu_development_2012}. In the plane tangential to the cortical surface the barrels constitute a somatotopic map of the whiskers, with cells of adjacent barrels responding most strongly and quickly to deflection of adjacent whiskers \citep{armstrong-james_flow_1992}. Barrel patterning reflects subcortical whisker maps comprising cell aggregates called barrelettes in the brainstem and barreloids in the thalamus.
%\cite{van_der_loos_barreloids_1976}.

Barrel formation requires afferent input from whisker stimulation and thalamic calcium waves \citep{anton-bolanos_prenatal_2019}, and depends on a complex network of axon guidance molecules such as ephrin-A5 and A7 and adhesion molecules such as cadherin-6 and 8 \citep{miller_epha7-ephrin-a5_2006}.
This network is orchestrated by interactions between morphogens Fgf8 and Fgf17 and transcription factors Emx2, Pax6, Sp8, and Coup-tf1 \citep{shimogori_fibroblast_2005}, which are expressed in gradients that mark orthogonal axes and can be manipulated to stretch, shrink, shift, and even duplicate barrels \cite{assimacopoulos_fibroblast_2012}.
% Removed ref: ,bishop_regulation_2000 after ..Coup-tf1

@@

\caption{\textbf{A.} Cortical whisker barrels from \cite{zheng_signal_2001}
  annotated with lines joining interstices into a putative Voronoi diagram
  ($\delta=0.022$). \textbf{B.}  Top: Stochastic initial condition for $a_i$
  with $i=1$ (barrel `$\alpha$'). The roll-off towards the boundary (thin
  line) ensures the boundary conditions are satisfied. Bottom: Guidance
  molecules $\rho_1(\mathbf{x})$ and $\rho_2(\mathbf{x})$ in the simulated
  cortex. $\rho_1(\mathbf{x})$ varies along the anterior-posterior axis;
  $\rho_2(\mathbf{x})$ is expressed along the lateral/medial axis, analogous
  to ephrin A5 \citep{vanderhaeghen_mapping_2000}. \textbf{C.}  Barreloid
  centres for mystacial whiskers taken from \cite{haidarliu_size_2001} are
  shown as black circles. The dashed grey line shows the boundary of the
  barreloid field. It is assumed that there are two linearly increasing,
  orthogonal signalling molecules in the thalamus, such that the location of
  each barreloid within the graded thalamic signal field can differentiate the
  efferent axons, with the result that axons from each barreloid interact to
  varying degrees with the cortical signalling gradients $\rho_1$ and
  $\rho_2$. Blue (green) arrows show, for each barreloid, the magnitude (and
  sign) of the interaction that its thalamocortical projection will have with
  $\rho_1$ ($\rho_2$). The mappings from barreloid location to interaction
  strength are linear. \textbf{D.}  Barrel cortex simulation with parameters
  $\alpha=3$, $\beta=20$, $k=3$, $D=0.2$, $\gamma\in\pm 2$ and
  $\delta{t}=0.0001$. Colours indicate $\arg\max_i(c_i)$ for 41 TC
  projections.  \textbf{E.}  The Honda Voronoi measure, $\delta$, (red) and a
  pattern fitting metric based on the distance, $d_i$, between simulated and
  experimental barrel centroids and the absolute difference in their areas,
  $\Delta A_i$ given by $\eta = \sum_i^N d_i^2 \times \sum_i^N \Delta A_i$
  (black, units L$^4$) plotted as a function of time for the simulations shown
  in D. \textbf{F.}  Guidance molecules in the simulated Fgf8 misexpression
  experiment. $\rho_1$ is mirrored over the anterior-posterior axis. The
  simulation indicates that the mirroring of the barrel field can occur as a
  result of the mirroring of guidance molecule 1.}
\label{fig:main}
\end{figure*}
